---
title: "Data Script Draft"
author: "Laura Gattis, Emmanuel Maduneme, Megan Denneny, Sam Lorenzo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true 
    toc_collapsed: true
    toc_depth: 4
    number_sections: true
    theme: lumen
bibliography: data_script_references.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      eval = TRUE,
                      message = FALSE)
library(tidyverse)
library(psych)
library(skimr)
library(readr)
library(janitor)
library(rio)
library(here)
library(ggthemes)
library(readxl)
library(DT)
library(tinytex)

```


## Introduction\centering
Laura and Megan making commits


## Methods \centering
We used the following packages: `janitor` firke2020janitor, `psych` revelle2015package, `skimr` @skimr, `here` @here, `knitr` @knitr, `rio` @rio, `readxl @readxl`,, `DT` @DT and `tidyverse` @tidyverse. 

After loading our data set **css_ghana**, we proceeded to use different functions like `filter()`, `mutate()`, `pibot_longer`, `pivot_wider`, `group_by`, and `select()` to tidy the data.

These are some descriptive information about the data. 

```{r tidying data}
css_ghana <- read_excel(here("Data", "Climate_Synthetic.xlsx"))



css_ghana <- css_ghana %>% 
  pivot_longer(c("Male", "Female"), 
               names_to = "Gender",
               values_to = "value") %>% 
  select(-(starts_with("1.")), -value,1:2,Gender, everything()) %>% 
  filter(Status == "Completed") %>% 
  rename(q8_cc_priority = "8. How should addressing issues of climate change be prioritized?") %>% 
  filter(!is.na(q8_cc_priority)) %>% 
  rename(age_generation =  "Which age generation below applies to you",
    q_10_cc_message_effec_providing_info  = 
      "10. Think about messages you’ve seen in the media about climate change. How effective were they in terms of providing information?",
    q_17_cc_infuence_child_desire = 
      "17. How much of an influence has climate change had on your desire to have children?",
    q7_cc_concerned = "7. How concerned are you about climate change?") %>% 
  mutate(age_gen_recoded = 
           case_when(age_generation ==  "Other" ~ "Others",
                     age_generation == "1997-2004" ~ "Gen_Z",
                     TRUE ~ "Millennials")) %>% 
   select(1:3, age_gen_recoded, everything())
 
names(css_ghana)[8] = "q_16_not_parent_chid_desire"
css_ghana <- css_ghana %>% 
  filter(!is.na(q_10_cc_message_effec_providing_info),
         !is.na( q_17_cc_infuence_child_desire),
         !is.na(q_16_not_parent_chid_desire),
         !is.na(q8_cc_priority))

css_ghana$Gender <- as.factor(css_ghana$Gender)

css_ghana$age_gen_recoded <- as.factor(css_ghana$age_gen_recoded)

 css_ghana$q7_cc_concerned <- recode(css_ghana$q7_cc_concerned, "Not at all" = 1, "Not very concerned" = 2 , "Slightly concerned" = 3, "Moderately concerned" = 4, "Very concerned" = 5,     
         "Extremely concerned" = 6)
  

## the row names were not properly written so I had to recode it  
css_ghana <- css_ghana %>% 
  mutate(q8_cc_priority = case_when(
    q8_cc_priority == "Low priority" ~"1",
    q8_cc_priority == "Moderate priority" ~ "3",
    q8_cc_priority == "Major priority" ~ "4",
    q8_cc_priority == "Top priority" ~ "5",
    q8_cc_priority == "Minor priority" ~ "2",
    q8_cc_priority == " Minor priority" ~ "2",
  ))

css_ghana$q_16_not_parent_chid_desire <- recode(css_ghana$q_16_not_parent_chid_desire,
  "Extremely high desire" = 7, "High desire" = 6,  "Minor desire" = 5, 
  "Moderate desire" = 4, "No desire or unsure" = 3,  "Somewhat dissatisfied" = 2,
  "Very low desire" = 1)  

css_ghana$q_17_cc_infuence_child_desire <- recode(css_ghana$q_17_cc_infuence_child_desire,
"Barely any influence" = 6, "Major influence" = 5,  "Minor influence" = 4,     "Moderate influence" = 3,
 "No influence at all" = 2, "Top influence"  = 1)


css_ghana$q_10_cc_message_effec_providing_info <- recode(css_ghana$q_10_cc_message_effec_providing_info,
"Extremely effective" = 6,               "Moderately effective"  = 5,                 
 "Not effective at all or Not applicable" = 4, "Slightly effective"  = 3,                 
 "Very effective" = 2 ,                    "Very low effect" = 1)


css_ghana %>% 
  select(!age_generation, !value)


## Pivot Wider for Assignment
css_ghana %>% 
  pivot_wider(
    names_from = age_gen_recoded,
    values_from = age_generation
  ) %>%
  pivot_longer(
    cols = c(Millennials, Gen_Z),
    names_to = "age_generation_pivot",
    values_to = "age_brackets"
  ) %>% 
  select(-starts_with("1.")) %>% 
  select(-value, -Others, -age_brackets) %>% 
  head() %>% 
  datatable()
  

```

Sam Making theese commits 
## Plotting those who were concerned about climate change by age generaton
```{r Descriptives }
names(css_ghana)

css_ghana %>% 
  ggplot(aes(q8_cc_priority))+
  geom_bar(aes(fill = age_gen_recoded), position = "dodge") +
  theme_minimal() +
  labs(x = "Climate Change Should Be Prioritized (1 = Low, 5 = High)",
         y = "Frequency",
         title = "Climate Change Priority By Age Generation",
       subtitle = "(Data obtained from a survey in Ghana, West Africa)")


```


```{r}
css_ghana %>% 
  mutate(q8_cc_priority = as.numeric(q8_cc_priority)) %>% 
 group_by(age_gen_recoded) %>% 
  summarize(mean_priority = mean(q8_cc_priority, na.rm = TRUE),
            mean_concerned = mean(q7_cc_concerned, na.rm = TRUE),
            mean_desire_for_child = mean(q_17_cc_infuence_child_desire, na.rm = TRUE)) %>% 
   ggplot(aes(age_gen_recoded, mean_desire_for_child, fill = age_gen_recoded)) +
  geom_col() +
  theme_minimal() +
   labs(x = "Age Generation",
         y = "Average Influence of Climate Change on Deire For Children",
         title = "Climate Change Influence on Child Desires",
         subtitle = "(Data obtained from a survey in Ghana, West Africa)") +
  scale_fill_manual(values = c("grey0",
                              "goldenrod", 
                              "goldenrod4")) +
  coord_flip()
 
```
```{r}
names(css_ghana)
css_ghana %>% 
  group_by(age_gen_recoded, q_10_cc_message_effec_providing_info) %>% 
  summarize(mean_message = mean(q_10_cc_message_effec_providing_info)) %>% 
  ggplot(aes(age_gen_recoded, mean_message)) + 
  geom_col(aes(fill = age_gen_recoded))
  
```


