---
title: "Rethinking Traditional Values: Analyzing How Messages of Climate Change Affect Family Planning Among Millennials and Generation Z"
author: "Laura Gattis, Emmanuel Maduneme, Megan Denneny, Sam Lorenzo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true 
    toc_collapsed: true
    toc_depth: 4
    number_sections: false
bibliography: data_script_references.bib
csl: apa.csl

---

# Abstract\centering

Climate change is perceived as a significant threat, especially among younger demographics. Millennials (born 1981-1996) are the most adamant about reversing the issue, with 71% stating it is a top priority to protect and sustain the planet for future generations. Generation Z adults, those over 18 (born 1997-2004), also share similar sentiments. According to prior studies, 1-in-4 U.S. adults under the age of 45 revealed that climate change risks motivate their decision to not have children. This study is the initial phase in a series of research. We have collected more than 150 surveys and began interviewing Millennials and Generation Z adults in Ghana. Both methods explore how participants view climate change messages and how this influences family planning. Our next step is to duplicate this study in the U.S. and eventually other countries to provide a comprehensive, comparative analysis of climate change and family planning perceptions across the globe.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      eval = TRUE,
                      message = FALSE)
library(tidyverse)
library(psych)
library(skimr)
library(readr)
library(janitor)
library(rio)
library(here)
library(ggthemes)
library(readxl)
library(DT)

```




# Introduction\centering

Climate change continues to be perceived as a significant risk among Americans, especially among younger demographics. Millennials, in particular, those born between 1981 and 1996, were found to be the group most adamant about prioritizing the issue of climate change, with 71 percent of Millennials indicating climate change to be the most important issue to address in order to protect and sustain the planet for future generations (Funk). According to a 2020 poll by Morning Consult, one-in-four adults under the age of 45 indicated that climate change plays a role (either major or minor) in their decision to not have children (Jenkins). A prior survey also measured the various reasons why young adults are choosing to have fewer than their ideal number of children, with 37 percent of respondents expressing a concern about global instability and 33 percent specifying the burden of climate change as key motives (Miller). 

We hypothesize that the way in which media messages about climate change are framed and interpreted negatively impacts Millennial and Generation Z perceptions about having children.

## Specific Aims/Study Objectives\centering
This analysis intends to answer the following research question: What effect do messages about climate change have on Millennials and Generation Z and their desire to have children?



# Methods, Materials, and Analysis\centering
We used `R programming` [@R] for analysis with the following packages: `janitor` [@janitor], `psych` [@psych], `skimr` [@skimr], `here` [@here], `knitr` [@knitr], `rio` [@rio], `readxl` [@readxl], `DT` [@DT] and `tidyverse` [@tidyverse]. 


After loading our data set **css_ghana**, we proceeded to use different functions like `filter()`, `mutate()`, `pivot_longer`, `pivot_wider`, `group_by`, and `select()` to tidy the data.


Researchers implemented a two-step, mixed-method approach to assess Millennial and Generation Z perceptions on family planning and to what extent (if any) climate change influences these views. For the purposes of this class, we will be analyzing the survey data we collected. The survey consisted of an online Likert-type questionnaire that was distributed to Ghanaian Millennials (those born between 1981 and 1996) and Generation Z adults (18 years old or older; those born between 1997 and 2004). The current sample size is made up of answers from 130 respondents. Through this survey, participants were asked questions about their personal stances on having children, perceptions of climate change, how much of an impact climate change has on their family planning, and their feelings towards the effectiveness of climate change messages. A scale from zero to five was provided for respondents to record their answers, with zero indicating no impact and five stipulating high impact. 

## Data Preparation Script Plan:

We are utilizing the data collected to quantitatively determine if there are any correlations between views amongst demographics and to see if responses tend to skew towards specific perceptions (e.g., high fear of climate change, low desire to have children). Our plan is to use our data to better visualize participant responses and categorize information into a variety of sets. We are employing a randomizer tool to skew the responses we received. 


These are some descriptive information about the data. 



```{r tidying data, include=FALSE}
css_ghana <- read_excel(here("Data", "Climate_Synthetic.xlsx"))



css_ghana <- css_ghana %>% 
  pivot_longer(c("Male", "Female"), 
               names_to = "Gender",
               values_to = "value") %>% 
  select(-(starts_with("1.")), -value,1:2,Gender, everything()) %>% 
  filter(Status == "Completed") %>% 
  rename(q8_cc_priority = "8. How should addressing issues of climate change be prioritized?") %>% 
  filter(!is.na(q8_cc_priority)) %>% 
  rename(age_generation =  "Which age generation below applies to you",
    q_10_cc_message_effec_providing_info  = 
      "10. Think about messages you’ve seen in the media about climate change. How effective were they in terms of providing information?",
    q_17_cc_infuence_child_desire = 
      "17. How much of an influence has climate change had on your desire to have children?",
    q7_cc_concerned = "7. How concerned are you about climate change?") %>% 
  mutate(age_gen_recoded = 
           case_when(age_generation ==  "Other" ~ "Others",
                     age_generation == "1997-2004" ~ "Gen_Z",
                     TRUE ~ "Millennials")) %>% 
   select(1:3, age_gen_recoded, everything())
 
names(css_ghana)[8] = "q_16_not_parent_chid_desire"
css_ghana <- css_ghana %>% 
  filter(!is.na(q_10_cc_message_effec_providing_info),
         !is.na( q_17_cc_infuence_child_desire),
         !is.na(q_16_not_parent_chid_desire),
         !is.na(q8_cc_priority))

css_ghana$Gender <- as.factor(css_ghana$Gender)

css_ghana$age_gen_recoded <- as.factor(css_ghana$age_gen_recoded)

 css_ghana$q7_cc_concerned <- recode(css_ghana$q7_cc_concerned, "Not at all" = 1, "Not very concerned" = 2 , "Slightly concerned" = 3, "Moderately concerned" = 4, "Very concerned" = 5,     
         "Extremely concerned" = 6)
  

## the row names were not properly written so I had to recode it  
css_ghana <- css_ghana %>% 
  mutate(q8_cc_priority = case_when(
    q8_cc_priority == "Low priority" ~"1",
    q8_cc_priority == "Moderate priority" ~ "3",
    q8_cc_priority == "Major priority" ~ "4",
    q8_cc_priority == "Top priority" ~ "5",
    q8_cc_priority == "Minor priority" ~ "2",
    q8_cc_priority == " Minor priority" ~ "2",
  ))

css_ghana$q_16_not_parent_chid_desire <- recode(css_ghana$q_16_not_parent_chid_desire,
  "Extremely high desire" = 7, "High desire" = 6,  "Minor desire" = 5, 
  "Moderate desire" = 4, "No desire or unsure" = 3,  "Somewhat dissatisfied" = 2,
  "Very low desire" = 1)  

css_ghana$q_17_cc_infuence_child_desire <- recode(css_ghana$q_17_cc_infuence_child_desire,
"Barely any influence" = 6, "Major influence" = 5,  "Minor influence" = 4,     "Moderate influence" = 3,
 "No influence at all" = 2, "Top influence"  = 1)


css_ghana$q_10_cc_message_effec_providing_info <- recode(css_ghana$q_10_cc_message_effec_providing_info,
"Extremely effective" = 6,               "Moderately effective"  = 5,                 
 "Not effective at all or Not applicable" = 4, "Slightly effective"  = 3,                 
 "Very effective" = 2 ,                    "Very low effect" = 1)


```



```{r}



## Pivot Wider for Assignment
css_ghana %>% 
  pivot_wider(
    names_from = age_gen_recoded,
    values_from = age_generation
  ) %>%
  pivot_longer(
    cols = c(Millennials, Gen_Z),
    names_to = "age_generation_pivot",
    values_to = "age_brackets"
  ) %>% 
  select(-starts_with("1.")) %>% 
  select(Gender, age_generation_pivot, everything(), -q_16_not_parent_chid_desire, -value, -Others,-age_brackets, -ID, -Status) %>% 
  datatable(class = 'cell-border stripe',
            colnames = c('Gender', 'Age Generation', 'Climate Change Concern', 'Climate Change Priority',
                         'Media Climate Change Message Efficacy', "Desire For Children"))
 
  

  
```


```{r Descriptives}
## Changed wording of title and legend slightly

css_ghana %>% 
  ggplot(aes(q8_cc_priority))+
  geom_bar(aes(fill = age_gen_recoded), position = "dodge") +
  theme_minimal() +
  labs(x = "Climate Change Should Be Prioritized (1 = Low, 5 = High)",
         y = "Frequency",
         title = "Climate Change Prioritization By Generation",
       subtitle = "(Data obtained from a survey in Ghana, West Africa)") +
  guides(fill=guide_legend(title= "Generation"))


```


```{r}
 ## Changing wording of title, subtitle, legend and x-axis slightly.

css_ghana %>% 
  mutate(q8_cc_priority = as.numeric(q8_cc_priority)) %>% 
 group_by(age_gen_recoded) %>% 
  summarize(mean_priority = mean(q8_cc_priority, na.rm = TRUE),
            mean_concerned = mean(q7_cc_concerned, na.rm = TRUE),
            mean_desire_for_child = mean(q_17_cc_infuence_child_desire, na.rm = TRUE)) %>% 
   ggplot(aes(age_gen_recoded, mean_desire_for_child, fill = age_gen_recoded)) +
  geom_col() +
  guides(fill=guide_legend(title= "Generation"))+
  theme_minimal() +
   labs(x = "Generation",
         y = "Average Influence of Climate Change on Desire to Have Children",
         title = "Climate Change Influence on Desire to Have Children",
         subtitle = "(Data obtained from a survey in Ghana, West Africa)") +
  scale_fill_manual(values = c("grey0",
                              "goldenrod", 
                              "goldenrod4")) +
    coord_flip()
   


```


```{r}

## Changing wording of title and y-axis slightly.

css_ghana %>% 
  group_by(age_gen_recoded) %>% 
  summarise(mean_message = mean(q_10_cc_message_effec_providing_info, na.rm = TRUE),
            avrg_desire = mean(q_17_cc_infuence_child_desire)) %>% 
  ggplot(aes(mean_message, avrg_desire)) + 
  geom_point(aes(col = age_gen_recoded), size = 5)+
  geom_smooth() +
   guides(fill=guide_legend(title= "Generation")) +
    labs(x = "Climate Change Media Effectiveness",
         y = "Average Influence of Climate Change on Desire to Have Children",
         title = "Climate Change Messaging on Desire to Have Children",
         subtitle = "(Data obtained from a survey in Ghana, West Africa)") 
   
  
```

# Initial Findings and Results\centering

As mentioned, we implemented a two-step, mixed-method approach to assess Millennial and Generation Z perceptions on family planning and to what extent climate change influences these views. The first step was to administer a Likert-type questionnaire where participants were asked about their personal stances on having children, perceptions of climate change, how much of an impact climate change has on their desire to have children, and their feelings towards the effectiveness of climate change messages. A scale from zero to five was provided for respondents to record their answers, with zero indicating no impact and five stipulating high impact. The survey concluded with an optional section where volunteers could indicate if they are willing to participate in a subsequent one-on-one interview to discuss their answers further. The survey data collected is currently being quantitatively assessed for correlations between views and to see if responses tend to skew towards specific perceptions (e.g., high fear of climate change, low desire to have children). The second part of the study is still ongoing, which consists of one-on-one interviews with volunteers. Upon completing data collection, we will transcribe and qualitatively code responses to observe if any common themes exist between participant responses.

Based on our preliminary analysis, Ghanaian participants who indicated that they were “extremely concerned” about climate change also tended to state that they have experienced some level of negative emotions when exposed to climate change messages, as well as mixed responses about their desire to have children due to climate change risks. Once we complete our analyses, we expect to uncover a more wholistic narrative of the Ghanaian perspective, which may include external influences such as cultural and societal values.


```{r findings}

## Added in two more variables and visual (below) that are a bit more comprehensive. 

research_questions <- css_ghana %>% 
  select(age_gen_recoded, q7_cc_concerned, q8_cc_priority, q_10_cc_message_effec_providing_info, q_17_cc_infuence_child_desire) %>% 
  rename(
    "How concerned are you about climate change?" = q7_cc_concerned,
    "How should addressing issues of climate change be prioritized?" = q8_cc_priority,
    "How effective is climate change news in terms of providing information?" = q_10_cc_message_effec_providing_info,
    "How much of an influence has climate change had on your desire to have children?" = q_17_cc_infuence_child_desire
    ) %>% 
  view()



```


```{r results}

research_questions %>% 
  pivot_longer(
    cols = -age_gen_recoded,
    names_to = "r_questions",
    values_to = "each_score",
    values_transform = list(each_score = as.integer)
  ) %>% 
  group_by(r_questions, age_gen_recoded) %>% 
  summarize(avrg_score = mean(each_score, na.rm = TRUE)) %>% 
 mutate(research_questions = factor(r_questions, level = c("How concerned are you about climate change?", "How should addressing issues of climate change be prioritized?", "How effective is climate change news in terms of providing information?", "How much of an influence has climate change had on your desire to have children?"))) %>% 
  ggplot(aes(x = age_gen_recoded, y = avrg_score, fill = age_gen_recoded)) + 
      geom_col(position = "dodge") + 
      facet_wrap(~research_questions, 
                 ncol = 1,
                 nrow = 6) + 
      theme_minimal()  + 
      scale_fill_colorblind() + 
      labs(
        title = "Climate Change and Family Planning Perceptions", 
        subtitle = "By Generation", 
        x = "Generation", 
        y = "Level of Perception",
        fill = "Generation", 
        caption = "(Data obtained from a survey in Ghana, West Africa)"
        )+
  coord_flip()

```
```{r}
css_ghana %>% 
 summarise(research_questions = mean(research_questions, na.rm = TRUE), 
           final_responses = mean(final_responses, na.rm = TRUE)) %>% 
  mutate(
    research_questions = c("How concerned are you about climate change?", "How should addressing issues of climate change be prioritized?", "Think about messages you’ve seen in the media about climate change. How effective were they in terms of providing information?", "How much of an influence has climate change had on your desire to have children?"),
    final_responses = factor(final_responses, level = c(1, 2, 3, 4, 5))
    ) %>% 
  ggplot(aes(x = NROW(css_ghana), y = final_responses, fill = age_gen_recoded)) + 
  geom_col(position = "dodge") + 
  facet_wrap(~research_questions, 
                  ncol = 1,
                  nrow = 6) + 
       theme_minimal()  + 
       scale_fill_colorblind() + 
       labs(
         title = "Climate Change and Family Planning Perceptions", 
         subtitle = "by Generation", 
         x = "Number of Responses", 
         y = "Level of Perception",
         fill = "Generation", 
         caption = "(Data obtained from a survey in Ghana, West Africa)"
       )
```

# Findings\centering
This graph demonstrates climate change and family perceptions by generation broken down by four key questions asked in interviews. 1.) How concerned are you about climate change? 2.) How should addressing issues of climate change be prioritized? 3.) How effective is climate change news in terms of providing information? 4.) How much of an influence has climate change had on your desire to have children? One key insight to highlight from the graph is that although Generation Z is the most concerned about climate change and addresses how these concerns should be prioritized, the data also shows that Generation Z is less likely to be influenced by climate change in their desire to have children. 

# Discussion and Broader impact\centering
Overall, our initial findings are impactful to better understanding the correlation between climate change and the lessened desire to have children amongst Millennials and Generation Z. The survey data collected in Ghana in 2022 will be helpful to compare to the U.S. and other countries in future data collection, all of which will come together to create a larger snapshot of this topic. Early findings suggest that climate change may have an impact on reproduction decision-making which can be insightful in locations like the U.S. where recent political and economical factors may also contribute. With this data, we can better infer how other countries may respond to similar structures and challenges as well as contribute solutions to current problems. 


\newpage
# References\centering
